---
title: "Customer Churn Analysis"
author: "Vignesh Venugopal"
date: "February 2, 2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("dplyr")
library("ggplot2")
```

## Data Summary

```{r, echo=F}
churnData <- read.csv("https://raw.githubusercontent.com/VickyVykciV/rProgramming/master/dataFiles/Cellphone.csv", header = T)
attach(churnData)
```
- Dimensions of Data
```{r, echo=FALSE}
dim(churnData)
```
- Variables of Data
```{r, echo=FALSE}
colnames(churnData)
```
- Structure of Data
```{r, echo=FALSE}
str(churnData)
```
- Summary
```{r, echo=FALSE}
summary(churnData)
```

## Data Preparation:

### Type Conversion:

After understanding the data, we see that few variables requires type conversion, because they have less levels to be called as continuous variables.

So the following variables are converted into **_factor_** variables

- Churn					
- ContractRenewal				
- DataPlan					
- CustServCalls

```{r, echo=FALSE}
churnData$Churn <- as.factor(churnData$Churn)
churnData$ContractRenewal <- as.factor(churnData$ContractRenewal)
churnData$DataPlan <- as.factor(churnData$DataPlan)
churnData$CustServCalls <- as.factor(churnData$CustServCalls)
```

## Exploratory Data Analysis:

### Univariate Analysis for Continuous Variables:

```{r, echo=FALSE}
par(mfrow = c(2, 2))
hist(AccountWeeks)
hist(DataUsage)
hist(DayMins)
hist(DayCalls)
hist(MonthlyCharge)
hist(OverageFee)
hist(RoamMins)
```

## Exploratory Data Analysis (cont..)

### Univariate Analysis for Categorical Variables:

```{r, echo=FALSE}
par(mfrow = c(2, 1))
barplot(table(ContractRenewal), xlab = 'Contract Renewal')
barplot(table(DataPlan), xlab = 'Data Plan')
barplot(table(CustServCalls), xlab = 'Customer Service Calls')
```

## Exploratory Data Analysis (cont..)

### Bivariate Analysis:

- Churning over Account Weeks

```{r, echo=FALSE}
churnOverAccountWeeks <-
  churnData %>% group_by(AccountWeeks, Churn) %>% summarise(ChurnCount = length(Churn)) %>%
  mutate(ChrunPerc = round((ChurnCount * 100 / sum(ChurnCount))), 0) %>%
  select(AccountWeeks, Churn, ChrunPerc)
ggplot(churnOverAccountWeeks, aes(x = AccountWeeks, y = ChrunPerc, fill = Churn)) + geom_bar(stat = "identity") + geom_text(aes(label = ChrunPerc),
            position = position_dodge(width = 0),
            vjust = 0)
```
