---
title: "Customer Churn Analysis"
author: "Vignesh Venugopal"
date: "February 2, 2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("dplyr")
library("ggplot2")
library("reshape2")
library("MASS")
library("dummies")
library("pscl")
library("caret")
library("pROC")
library("lmtest")
library("corrplot")
library("Deducer")
source("https://raw.githubusercontent.com/VickyVykciV/rProgramming/master/functions/woe.R")
source("https://raw.githubusercontent.com/VickyVykciV/rProgramming/master/functions/deciling.R")
```

## Data Summary

```{r, echo=F}
cost <- function(theta)
{
m <- nrow(Predictor_Var)
g <- sigmoid(Predictor_Var%*%theta)
J<-(1/m)*sum((-Y*log(g)) - ((1-Y)*log(1-g)))
return(J)
}

churnData <- read.csv("https://raw.githubusercontent.com/VickyVykciV/rProgramming/master/dataFiles/Cellphone.csv", header = T)
attach(churnData)
```
- Dimensions of Data
```{r, echo=FALSE}
dim(churnData)
```
- Variables of Data
```{r, echo=FALSE}
colnames(churnData)
```
- Structure of Data
```{r, echo=FALSE}
str(churnData)
```
- Summary
```{r, echo=FALSE}
summary(churnData)
```

## Data Preparation:

### Type Conversion:

After understanding the data, we see that few variables requires type conversion, because they have less levels to be called as continuous variables.

So the following variables are converted into **_factor_** variables

- Churn					
- ContractRenewal				
- DataPlan

```{r, echo=FALSE}
churnData$Churn <- as.factor(churnData$Churn)
churnData$ContractRenewal <- as.factor(churnData$ContractRenewal)
churnData$DataPlan <- as.factor(churnData$DataPlan)

str(churnData)
```

## Exploratory Data Analysis:

### Univariate Analysis for Continuous Variables:

```{r, echo=FALSE}
par(mfrow = c(2, 5))
hist(AccountWeeks)
barplot(table(ContractRenewal), xlab = 'Contract Renewal', main = 'Bar Plot of Contract Renewal')
barplot(table(DataPlan), xlab = 'Data Plan', main = 'Bar Plot of Data Plan')
hist(DataUsage)
hist(CustServCalls)
hist(DayMins)
hist(DayCalls)
hist(MonthlyCharge)
hist(OverageFee)
hist(RoamMins)
```


## Exploratory Data Analysis (cont..)

### Bivariate Analysis:

- Churning over Account Weeks

```{r, echo=FALSE}
churnOverAccountWeeks <-
  churnData %>% group_by(Churn) %>% summarise(Avg_Tenure_Contract = round(mean(AccountWeeks), 0))

ggplot(churnOverAccountWeeks, aes(x = Churn, y = Avg_Tenure_Contract)) + geom_bar(stat = "identity") + geom_text(aes(label = Avg_Tenure_Contract),
            position = position_dodge(width = 0),
            vjust = 0)
```



```{r, echo=F}
Contract_Tb<-  churnData %>% group_by(Churn,ContractRenewal) %>% summarise(Percentage = length(ContractRenewal)/nrow(churnData))
ggplot(Contract_Tb, aes(x= Churn, y=Percentage)) + 
  geom_bar(aes(fill = ContractRenewal), stat="identity")+
  facet_grid(~ContractRenewal)
```



```{r, echo=F}
dataPlan_Tb<-  churnData %>% group_by(Churn,DataPlan) %>% summarise(Perc_Data_Opt = length(DataPlan)/nrow(churnData))
ggplot(dataPlan_Tb, aes(x= DataPlan, y=Perc_Data_Opt)) + 
  geom_bar(aes(fill = Churn), stat="identity") +
  facet_grid(~Churn)
```



```{r, echo=F}
churnOverDataUsage <-
  churnData %>% group_by(Churn) %>% summarise(Avg_Data_Usage = mean(DataUsage))
ggplot(churnOverDataUsage, aes(x = Churn, y = Avg_Data_Usage)) + geom_bar(stat = "identity") + geom_text(aes(label = Avg_Data_Usage),
            position = position_dodge(width = 0),
            vjust = 0)
```



```{r, echo=F}
custCalls_Tb<-  churnData %>% group_by(Churn,CustServCalls) %>% summarise(Perc_Service_Calls = length(CustServCalls)/nrow(churnData))
ggplot(custCalls_Tb, aes(x= CustServCalls, y=Perc_Service_Calls)) + 
  geom_bar(aes(fill = Churn), stat="identity") +
  facet_grid(~Churn)
```



```{r, echo=F}
churnOverDayMins <-
  churnData %>% group_by(Churn) %>% summarise(Avg_Day_Mins = round(mean(DayMins), 0))
ggplot(churnOverDayMins, aes(x = Churn, y = Avg_Day_Mins )) + geom_bar(stat = "identity") + geom_text(aes(label = Avg_Day_Mins),
            position = position_dodge(width = 0),
            vjust = 0)
```



```{r, echo=F}
churnOverDayCalls <-
  churnData %>% group_by(Churn) %>% summarise(Avg_Day_Call = round(mean(DayCalls), 0))
ggplot(churnOverDayCalls, aes(x = Churn, y = Avg_Day_Call )) + geom_bar(stat = "identity") + geom_text(aes(label = Avg_Day_Call),
            position = position_dodge(width = 0),
            vjust = 0)
```



```{r, echo=F}
churnOverMonthlyCharges <-
  churnData %>% group_by(Churn) %>% summarise(charge = round(mean(MonthlyCharge), 0))
ggplot(churnOverMonthlyCharges, aes(x = Churn, y = charge )) + geom_bar(stat = "identity") + geom_text(aes(label = charge),
            position = position_dodge(width = 0),
            vjust = 0)
```



```{r, echo=F}
churnOverOverUsageFee <-
  churnData %>% group_by(Churn) %>% summarise(fee = round(mean(OverageFee), 0))
ggplot(churnOverOverUsageFee, aes(x = Churn, y = fee )) + geom_bar(stat = "identity") + geom_text(aes(label = fee),
            position = position_dodge(width = 0),
            vjust = 0)
```



```{r, echo=F}
churnOverRoaming <- 
  churnData %>% group_by(Churn) %>% summarise(roaming = round(mean(RoamMins), 0))
ggplot(churnOverRoaming, aes(x = Churn, y = roaming )) + geom_bar(stat = "identity") + geom_text(aes(label = roaming),
            position = position_dodge(width = 0),
            vjust = 0)
```

## Hypothesis Testing

```{r, echo=F}
t.test(AccountWeeks~Churn)
t.test(DataUsage~Churn)
t.test(DayMins~Churn)
t.test(DayCalls~Churn)
t.test(MonthlyCharge~Churn)
t.test(OverageFee~Churn)
t.test(RoamMins~Churn)
```

### Importane of Varaible
```{r, echo=F}
categoricalVaraible <- c("ContractRenewal", "DataPlan")
CategoricalFeatures <- do.call('rbind', lapply(categoricalVaraible, WoE, data = churnData, 'Churn', F))
(CategoricalFeatures <- CategoricalFeatures[order(-CategoricalFeatures$IV),])
```

##Data Sampling:

```{r, echo=F}
set.seed(123)
smp_size <- floor(0.70 * nrow(churnData))
train_ind <- sample(seq_len(nrow(churnData)), size = smp_size)
trainData <- churnData[train_ind, ]
testData <- churnData[-train_ind, ]
```

### Complete Data
```{r, echo=F}
nrow(churnData)
prop.table(table(churnData$Churn))
```

### Development Sample
```{r, echo=F}
nrow(trainData)
prop.table(table(trainData$Churn))
```

### Holdout Sample
```{r, echo=F}
nrow(testData)
prop.table(table(testData$Churn))
```
## Model Building

### Model 1:

```{r, echo=F}
model1 <- glm(Churn~., family = "binomial", data = trainData)
summary(model1)
```

## Model Building(cont..)

### Stepwise
```{r, echo=F}
stepAIC(model1)
```

## Model Building(cont..)
### Model 2
```{r, echo=F}
model2 <- glm(formula = Churn ~ ContractRenewal + DataPlan + CustServCalls + 
    DayMins + OverageFee + RoamMins, family = "binomial", data = trainData)
summary(model2)
```

## Model Statistics

### Maximum Likelyhood Test
```{r, echo=F}
lrtest(model2)
```

### Pseudo R-Square
```{r, echo=F}
pR2(model2)
```

### Confusion Matrix
```{r, echo=F}
trainData$Prediction <- predict(model2, newdata = trainData, type = "response")
trainData$Prediction_Class <- ifelse(trainData$Prediction < 0.5, 0, 1)
confusionMatrix(trainData$Churn, trainData$Prediction_Class, positive = "1")
```

## ODDS Ratio

```{r, echo=F}
exp(cbind('Odds Ratio' = coef(model2), confint(model2)))
```

## Correlation

```{r, echo=F}
corMatrix <- cor(trainData[, -c(1, 3, 4, 12, 13)])
corrplot(corMatrix, method = "circle")
```

## Validating the model
```{r, echo=F}
testData$Prediction <- predict(model2, newdata = testData, type = "response")
testData$Prediction_Class <- ifelse(testData$Prediction < 0.4, 0, 1)
confusionMatrix(testData$Churn, testData$Prediction_Class)
```

## ROC Curve

```{r, echo=F}
# Train
p <- predict(model2, newdata=trainData, type="response")
pr <- prediction(p, trainData$Churn)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc
#Test
p <- predict(model2, newdata=testData, type="response")
pr <- prediction(p, testData$Churn)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc
```

```{r, echo=F}
trainData$deciles <- decile(trainData$Prediction)
KS_Table <- data.frame()
for (i in 1:10)
{
  Cnt = length(trainData$Churn[trainData$deciles  == i])
  Cnt_Resp = sum(as.numeric(trainData$Churn[trainData$deciles  == i] == 1))
  Cnt_Non_Resp = sum(as.numeric(trainData$Churn[trainData$deciles  == i] == 0))
  KS_Tab <- cbind(i, Cnt, Cnt_Resp, Cnt_Non_Resp)
  KS_Table <- rbind(KS_Table, KS_Tab)
}

KS_Table$Resp_Rate <-  KS_Table$Cnt_Resp * 100 / KS_Table$Cnt
KS_Table$Cum_Resp <- cumsum(KS_Table$Cnt_Resp)
KS_Table$Cum_Non_Resp <- cumsum(KS_Table$Cnt_Non_Resp)
KS_Table$Cum_Rel_Resp <- KS_Table$Cum_Resp / sum(KS_Table$Cnt_Resp)
KS_Table$Cum_Rel_Non_Resp <- KS_Table$Cum_Non_Resp / sum(KS_Table$Cnt_Non_Resp)
KS_Table$KS <- abs(KS_Table$Cum_Rel_Resp - KS_Table$Cum_Rel_Non_Resp)
KS_Table 
write.csv(KS_Table,"KS1.csv")

#Deciling Function 
#Test

testData$deciles <- decile(testData$Prediction)
KS_Table <- data.frame()
for (i in 1:10)
{
  Cnt = length(testData$Churn[testData$deciles  == i])
  Cnt_Resp = sum(as.numeric(testData$Churn[testData$deciles  == i] == 1))
  Cnt_Non_Resp = sum(as.numeric(testData$Churn[testData$deciles  == i] == 0))
  KS_Tab <- cbind(i, Cnt, Cnt_Resp, Cnt_Non_Resp)
  KS_Table <- rbind(KS_Table, KS_Tab)
}

KS_Table$Resp_Rate <-  KS_Table$Cnt_Resp * 100 / KS_Table$Cnt
KS_Table$Cum_Resp <- cumsum(KS_Table$Cnt_Resp)
KS_Table$Cum_Non_Resp <- cumsum(KS_Table$Cnt_Non_Resp)
KS_Table$Cum_Rel_Resp <- KS_Table$Cum_Resp / sum(KS_Table$Cnt_Resp)
KS_Table$Cum_Rel_Non_Resp <- KS_Table$Cum_Non_Resp / sum(KS_Table$Cnt_Non_Resp)
KS_Table$KS <- abs(KS_Table$Cum_Rel_Resp - KS_Table$Cum_Rel_Non_Resp)
KS_Table 
```
