---
title: "Customer Churn Analysis"
author: "Vignesh Venugopal"
date: "February 2, 2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("dplyr")
library("ggplot2")
library("reshape2")
library("MASS")
library("dummies")
library("pscl")
library("caret")
library("pROC")
library("lmtest")
source("https://raw.githubusercontent.com/VickyVykciV/rProgramming/master/functions/woe.R")
```

## Data Summary

```{r, echo=F}
churnData <- read.csv("https://raw.githubusercontent.com/VickyVykciV/rProgramming/master/dataFiles/Cellphone.csv", header = T)
attach(churnData)
```
- Dimensions of Data
```{r, echo=FALSE}
dim(churnData)
```
- Variables of Data
```{r, echo=FALSE}
colnames(churnData)
```
- Structure of Data
```{r, echo=FALSE}
str(churnData)

head(churnData)
```
- Summary
```{r, echo=FALSE}
summary(churnData)
```

## Data Preparation:

### Type Conversion:

After understanding the data, we see that few variables requires type conversion, because they have less levels to be called as continuous variables.

So the following variables are converted into **_factor_** variables

- Churn					
- ContractRenewal				
- DataPlan					
- CustServCalls

```{r, echo=FALSE}
churnData$Churn <- as.factor(churnData$Churn)
churnData$ContractRenewal <- as.factor(churnData$ContractRenewal)
churnData$DataPlan <- as.factor(churnData$DataPlan)

str(churnData)
```

## Exploratory Data Analysis:

### Univariate Analysis for Continuous Variables:

```{r, echo=FALSE}
par(mfrow = c(2, 3))
hist(AccountWeeks)
barplot(table(ContractRenewal), xlab = 'Contract Renewal')
barplot(table(DataPlan), xlab = 'Data Plan')
hist(DataUsage)
hist(CustServCalls)
hist(DayMins)
hist(DayCalls)
hist(MonthlyCharge)
hist(OverageFee)
hist(RoamMins)
```


## Exploratory Data Analysis (cont..)

### Bivariate Analysis:

- Churning over Account Weeks

```{r, echo=FALSE}
churnOverAccountWeeks <-
  churnData %>% group_by(Churn) %>% summarise(Avg_Tenure_Contract = round(mean(AccountWeeks), 0))

ggplot(churnOverAccountWeeks, aes(x = Churn, y = Avg_Tenure_Contract)) + geom_bar(stat = "identity") + geom_text(aes(label = Avg_Tenure_Contract),
            position = position_dodge(width = 0),
            vjust = 0)

Contract_Tb<-  churnData %>% group_by(Churn,ContractRenewal) %>% summarise(Per_Renewal = length(ContractRenewal)/nrow(churnData))
ggplot(Contract_Tb, aes(x= ContractRenewal, y=Per_Renewal)) + 
  geom_bar(aes(fill = Churn), stat="identity") +
  facet_grid(~Churn)

dataPlan_Tb<-  churnData %>% group_by(Churn,DataPlan) %>% summarise(Perc_Data_Opt = length(DataPlan)/nrow(churnData))
ggplot(dataPlan_Tb, aes(x= DataPlan, y=Perc_Data_Opt)) + 
  geom_bar(aes(fill = Churn), stat="identity") +
  facet_grid(~Churn)

churnOverDataUsage <-
  churnData %>% group_by(Churn) %>% summarise(Avg_Data_Usage = mean(DataUsage))
ggplot(churnOverDataUsage, aes(x = Churn, y = Avg_Data_Usage)) + geom_bar(stat = "identity") + geom_text(aes(label = Avg_Data_Usage),
            position = position_dodge(width = 0),
            vjust = 0)

custCalls_Tb<-  churnData %>% group_by(Churn,CustServCalls) %>% summarise(Perc_Service_Calls = length(CustServCalls)/nrow(churnData))
ggplot(custCalls_Tb, aes(x= CustServCalls, y=Perc_Service_Calls)) + 
  geom_bar(aes(fill = Churn), stat="identity") +
  facet_grid(~Churn)

churnOverDayMins <-
  churnData %>% group_by(Churn) %>% summarise(Avg_Day_Mins = round(mean(DayMins), 0))
ggplot(churnOverDayMins, aes(x = Churn, y = Avg_Day_Mins )) + geom_bar(stat = "identity") + geom_text(aes(label = Avg_Day_Mins),
            position = position_dodge(width = 0),
            vjust = 0)

churnOverDayCalls <-
  churnData %>% group_by(Churn) %>% summarise(Avg_Day_Call = round(mean(DayCalls), 0))
ggplot(churnOverDayCalls, aes(x = Churn, y = Avg_Day_Call )) + geom_bar(stat = "identity") + geom_text(aes(label = Avg_Day_Call),
            position = position_dodge(width = 0),
            vjust = 0)

churnOverMonthlyCharges <-
  churnData %>% group_by(Churn) %>% summarise(charge = round(mean(MonthlyCharge), 0))
ggplot(churnOverMonthlyCharges, aes(x = Churn, y = charge )) + geom_bar(stat = "identity") + geom_text(aes(label = charge),
            position = position_dodge(width = 0),
            vjust = 0)

churnOverOverUsageFee <-
  churnData %>% group_by(Churn) %>% summarise(fee = round(mean(OverageFee), 0))
ggplot(churnOverOverUsageFee, aes(x = Churn, y = fee )) + geom_bar(stat = "identity") + geom_text(aes(label = fee),
            position = position_dodge(width = 0),
            vjust = 0)

churnOverRoaming <- 
  churnData %>% group_by(Churn) %>% summarise(roaming = round(mean(RoamMins), 0))
ggplot(churnOverRoaming, aes(x = Churn, y = roaming )) + geom_bar(stat = "identity") + geom_text(aes(label = roaming),
            position = position_dodge(width = 0),
            vjust = 0)
```

## Hypothesis Testing

```{r, echo=F}
t.test(AccountWeeks~Churn)
t.test(DataUsage~Churn)
t.test(DayMins~Churn)
t.test(DayCalls~Churn)
t.test(MonthlyCharge~Churn)
t.test(OverageFee~Churn)
t.test(RoamMins~Churn)
```

### Importane of Varaible
```{r, echo=F}
categoricalVaraible <- c("ContractRenewal", "DataPlan")
CategoricalFeatures <- do.call('rbind', lapply(categoricalVaraible, WoE, data = churnData, 'Churn', F))
(CategoricalFeatures <- CategoricalFeatures[order(-CategoricalFeatures$IV),])
```

## Model

### Model 1:

```{r, echo=F}

set.seed(1)
smp_size <- floor(0.70 * nrow(churnData))
train_ind <- sample(seq_len(nrow(churnData)), size = smp_size)
trainData <- churnData[train_ind, ]
testData <- churnData[-train_ind, ]
nrow(churnData)
prop.table(table(churnData$Churn))
nrow(trainData)
prop.table(table(trainData$Churn))
nrow(testData)
prop.table(table(testData$Churn))

model1 <- glm(Churn~., family = "binomial", data = trainData)
summary(model1)
stepAIC(model1)

model2 <- glm(formula = Churn ~ ContractRenewal + DataPlan + CustServCalls + 
    DayMins + OverageFee + RoamMins, family = "binomial", data = trainData)
summary(model2)

lrtest(model2)

pR2(model2)

trainData$Prediction <- predict(model2, newdata = trainData, type = "response")
trainData$Prediction_Class <- ifelse(trainData$Prediction < 0.5, 0, 1)

##Misclassification Error

cm <- confusionMatrix(trainData$Churn, trainData$Prediction_Class)

cm
```
