---
title: "Grocery Store Recommendation Engine"
author: "Group 7"
date: "April 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(arules)
library(arulesViz)
```

##Reading the data
```{r, echo=F}
transactions = read.transactions("https://raw.githubusercontent.com/vigneshvenugopal/rProgramming/master/dataFiles/GroceryStore_Transaction_Data.csv", 
                            format = "single", sep = ",",
                          cols = c("Transaction", "Product"), rm.duplicates = T)
```

##Inspect Data
```{r, echo=F}
inspect(head(transactions))
summary(transactions)
```
It is noted that 72% of the transactions have only one item. 17% of the transactions have 2 items. 
The matrix of items against transactions is sparse with a density of 0.08. 
In the given dataset, a transaction contains only 1.43 items on an average.


##Transaction Plots
### Top 10, most frequent items

```{r, echo=F}
itemFrequencyPlot(transactions, topN=10)
itemFrequencyPlot(transactions, support=0.01, topN=10)
```

The top 10 items, with the support of 1% are

Magazine, Candy Bar, Toothpaste, Greeting Cards, Pens, Pencils, Perfume, Toothbrush, Wrapping Paper & Photo Processing

### Running apriori rules on the above mentioned items with a confidence of 10%

To avoid single item recommendations that are not very useful, we specify minlen = 2

```{r}
rule <- apriori(data=transactions, parameter=list(supp=0.01,conf = 0.1, minlen=2),
                           appearance = list(default="lhs",
                           both = c("Magazine", "Candy Bar", "Toothpaste","Greeting Cards",
                                    "Pens","Pencils","Perfume","Toothbrush","Wrapping Paper",
                                    "Photo Processing")),
                           control = list(verbose=T))
```

###Sorting the rules based on the LIFT attained

```{r}
inspect(sort(rule, by="lift"))
```
There are 49 item sets with minimum support value of 1% and minimum confidence level of 10%

### Understanding association rules
A common approach is to take the result of learning association rules and divide them into three categories

Actionable - The goal of a market basket analysis is to find actionable associations, or rules that provide a clear and useful insight. Some rules are clear, others are useful; it is less common to find a combination of both of these factors.

Trivial - Any rules that are so obvious that they are not worth mentioning, they are clear, but not useful.

Inexplicable - If the connection between the items is so unclear that figuring out how to use the information for action would require additional research


### Deeper Look
```{r}
inspect(sort(rule, by = "lift")[1:10])
```


The first rule, with a lift of 3.06, implies that people who buy Perfumes are nearly three times more likely to buy Toothbrush than the typical customer and vice-versa also stands true

### Recommendataions
Item set with LIFT atleast 1, only should be recommended

Based on the RHS item, the item set combination which gives the highest LIFT is recommended, please see the example below mentioned for Toothbrush and Candy Bar

###Subset of Assocaition Rule

```{r}
ruleSubset = subset(rule, items %in% c("Toothbrush","Candy Bar"))
inspect(sort(ruleSubset, by = "lift")[1:10])
```

### Association rules - Scatter Plot
```{r, echo=F, message=F}
plot(rule)
```