---
title: "Grocery Store Recommendation Engine"
author: "Group 7"
date: "April 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(arules)
library(arulesViz)
```

##Reading the data
```{r, echo=F}
data <- read.csv("https://raw.githubusercontent.com/vigneshvenugopal/rProgramming/master/dataFiles/GroceryStore_Transaction_Data.csv", header = T, sep = ",", colClasses = c('character', rep('factor', 4)))
data <-  data[, -1]
str(data)
head(data)
```

##Converting the data into transaction data
Since the data is in Basket Format, we have to convert into Single transaction data
```{r, echo=F, warning=F}
transactions <- as(split(rep(data[,"Product"], data[,"Quantity"]), data[,"Transaction"]), "transactions")
```

##Sample Transaction Data
```{r, echo=F}
inspect(head(transactions, 1))
```
##Transaction and Item counts
```{r, echo=F}
transactions
```
##Transaction Plots

Displaying the top 10, most frequent items

```{r, echo=F}
itemFrequencyPlot(transactions, topN = 10, type = 'absolute')
```

##Finding the top 10 items sets, which has a minimum support of 1%

```{r, echo=F, message=F}

itemFrequency(transactions)

rules_1 <- apriori(transactions, parameter = list(supp = 0.01))
top.support <- sort(rules_1, decreasing = TRUE, na.last = NA, by = "support")
inspect(head(top.support, 10))
```

##Of the above 10 items-sets, finding the items-sets with minimum confidence of 10%

```{r, echo=F, message=F}
rules_2 <- apriori(transactions, parameter = list(supp = 0.01, conf = 0.1))
top.support <- sort(rules_2, decreasing = TRUE, na.last = NA, by = c("support", "confidence"))
inspect(head(top.support, 10))
```
